Patient-Specific Neural Field Model

### 2010-07-19 ltx/data_paper/DataPaper.tex
added to the model derivation and the pre-processing steps

where $T_s$ is the time step or sampling period. Rearranging equation~\ref{EulerMethod} gives the integro-difference equation (IDE) form

@@

	v_{t+T_s}\left(\mathbf{r}\right) = 

@@

To reduce the model to a finite dimensional system the neural field is approximated by a basis function decomposition where

@@

	\label{DefFieldDecomp} v_t\left(\mathbf{r}\right) \approx \boldsymbol{\phi}^{\top}\left(\mathbf{r}\right) \mathbf{x}_t, 

@@

where $\mathbf{x}_t$ is a state vector that weights another vector of two-dimension Gaussian basis functions, $\boldsymbol{\phi}(\mathbf{r})$, described by

@@

	\boldsymbol{\phi}^{\top}(\mathbf{r})\mathbf{x}_{t+1}= T_s\int_\Omega{f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t )w(\mathbf{r}-\mathbf{r}')\textrm{d}\mathbf{r}'}

@@

 	\int_\Omega {\boldsymbol{\phi} \left(\mathbf{r}\right)\boldsymbol{\phi}^{\top}\left(\mathbf{r}\right) \textrm{d}\mathbf{r}} \mathbf{x}_{t+1} = T_s \int_\Omega {\boldsymbol{\phi} (\mathbf{r}) \int_\Omega {w(\mathbf{r}-\mathbf{r}') f(\boldsymbol{\phi}^{\top}(\mathbf{r}') \mathbf{x}_t ) \textrm{d}\mathbf{r}'}\textrm{d}\mathbf{r}} + \xi\int_\Omega{\boldsymbol{\phi}(\mathbf{r})\boldsymbol{\phi}^{\top}(\mathbf{r})\textrm{d}\mathbf{r}} \mathbf{x}_t + \int_\Omega{\boldsymbol{\phi} (\mathbf{r}) e_t(\mathbf{r})\textrm{d}\mathbf{r}}. 

@@

enables the pseudo inverse of $\boldsymbol{\phi(\mathbf{r})}$ to be taken such that the state vector can is isolated on the left-hand-side as

@@

	 \int_\Omega w(\mathbf{r}-\mathbf{r}')f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) \textrm{d}\mathbf{r}' \textrm{d}\mathbf{r} 

@@

@@

\begin{itemize}
	\item Show medical imaging
	\item Show LFP Data (background, seizure, post-seizure)
	\item Show spectral properties of LFPs
\end{itemize}

@@

The pre-processing steps consisted of:
\begin{itemize}
	\item Identification and removal (set to nan) of severely noise corrupted channels (16 of 96 channels);
	\item Implementation of a common average reference (CAR) montage; 
	\item Low-pass filter data, with cut-off at 100~Hz;
	\item Decimate data from 30~kHz to 5~kHz temporal sampling;
	\item Spatially interpolate missing data to facilitate spatial frequency analysis.
\end{itemize}

\subsection{Spatiotemporal Data Analysis}
The first 5 channels of the LFP time series from the seizure is shown in Figure~\ref{fig:TimeSeries}. For the analysis, the time series was segmented into three periods; the pre-seizure period, seizure period and post-seizure period. The seizure onset is marked by the red dotted line and the seizure end is marked by the blue dashed line.

\begin{figure}[!ht]
\begin{center}
\includegraphics{./Figures/LFPs.eps}
\end{center}
\caption{{\bf Local field potentials from the micro-electrode array}. Data recorded from the first five channels of the array. The red dotted line indicates the seizure onset and the blue dashed line indicates the seizure end.}
\label{fig:TimeSeries}
\end{figure}

>>>>

where $T_s$ is the time step or sampling period. To simplify the notation, the sample at the the current time step shall be indexed by $t$ and the future time step by $t+1$ for the rest of the paper. Rearranging equation~\ref{EulerMethod} gives the integro-difference equation (IDE) form

@@

	v_{t+1}\left(\mathbf{r}\right) = 

@@

To reduce the model to a finite-dimensional system the neural field is approximated by a basis function decomposition where

@@

	\label{DefFieldDecomp} v_t\left(\mathbf{r}\right) \approx \boldsymbol{\phi}^{\top}\left(\mathbf{r}\right) \mathbf{x}_t 

@@

and $\mathbf{x}_t$ is a state vector that weights a vector of two-dimension Gaussian basis functions, $\boldsymbol{\phi}(\mathbf{r})$, described by

@@

	\boldsymbol{\phi}^{\top}(\mathbf{r})\mathbf{x}_{t+1}= T_s\int_\Omega{f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t )w(\mathbf{r},\mathbf{r}')\textrm{d}\mathbf{r}'}

@@

 	\int_\Omega {\boldsymbol{\phi} \left(\mathbf{r}\right)\boldsymbol{\phi}^{\top}\left(\mathbf{r}\right) \textrm{d}\mathbf{r}} \mathbf{x}_{t+1} = T_s \int_\Omega {\boldsymbol{\phi} (\mathbf{r}) \int_\Omega {w(\mathbf{r},\mathbf{r}') f(\boldsymbol{\phi}^{\top}(\mathbf{r}') \mathbf{x}_t ) \textrm{d}\mathbf{r}'}\textrm{d}\mathbf{r}} + \xi\int_\Omega{\boldsymbol{\phi}(\mathbf{r})\boldsymbol{\phi}^{\top}(\mathbf{r})\textrm{d}\mathbf{r}} \mathbf{x}_t + \int_\Omega{\boldsymbol{\phi} (\mathbf{r}) e_t(\mathbf{r})\textrm{d}\mathbf{r}}. 

@@

enables the pseudo inverse of $\boldsymbol{\phi(\mathbf{r})}$ to be taken, such that the state vector can is isolated on the left-hand-side as

@@

	 \int_\Omega w(\mathbf{r},\mathbf{r}')f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) \textrm{d}\mathbf{r}' \textrm{d}\mathbf{r} 

@@

A property of the Gaussian basis function decomposition it that the Gaussianality of the disturbance term is maintained. The decomposition disturbance term is defined as
\begin{equation}\label{eq:Wt} 
	\mathbf{e}_t \triangleq \boldsymbol{\Gamma}^{-1}\int_\Omega {\boldsymbol{\phi} ( \mathbf{r} )e_t( \mathbf{r} )\textrm{d}\mathbf{r}},
where $\mathbf{e}_t \sim\mathcal{N}(\mathbf 0,\boldsymbol\Sigma_e)$. The covariance matrix is defined as (see the Supporting Information, S1, for the derivation)
	\boldsymbol\Sigma_e \triangleq \mathbf{\Gamma}^{-1}\int_{\Omega}\int_{\Omega}\boldsymbol{\phi}\left(\mathbf r\right) \gamma\left(\mathbf r- \mathbf r' \right)\boldsymbol{\phi}\left(\mathbf r'\right)^{\top}d\mathbf r' d\mathbf r\mathbf{\Gamma}^{- \top}. 

@@

@@

The first pre-processing step was to identify channels that had very poor signal quality. These channels were set to the $nan$ data type. Following this, the data was re-referenced to a common average montage to remove common mode artifact and the effect of the reference electrode. Next the data was low-pass filtered with a cut-off frequency of 100~Hz. The data was then resampled from 30~kHz to 5~kHz. To enable spatial visualisation and spatial frequency analysis, the corrupted channels, set to $nan$, were interpolated using all neighbouring channels.

\subsection{Spatiotemporal Data Analysis}
The first 8 channels of the LFP time series from the seizure is shown in Figure~\ref{fig:TimeSeries}. For the analysis, the time series was segmented into three periods; the pre-seizure period, seizure period and post-seizure period. The seizure onset is marked by the red dotted line and the seizure end is marked by the blue dashed line.

\begin{figure}[!ht]
\begin{center}
\includegraphics{./Figures/LFPs.eps}
\end{center}
\caption{{\bf Local field potentials from the micro-electrode array}. Data recorded from the first 8 channels of the array. The red dotted line indicates the seizure onset and the blue dashed line indicates the seizure end.}
\label{fig:TimeSeries}
\end{figure}
